GRN inference in multiple species

### 2015-04-29 paper1.tex
hello

\section{Results}

\subsection{Fusing similar network has the same effect as increasing the number of experimental conditions}
When the sample size of gene expression increases, the performance

\subsection{SCAD successfully identifies and unfuses 'neofunctionalized' genes}

>>>>

\section{Results}
We tested the ability of fused regression to improve network inference performance, as well as the ability of adaptive-fusion to learn the conservation of gene interactions. We measured network inference performance with the area under the precision recall curve (AUPR).

\subsection{Network similarity and fusion performance}
We assessed how varying the similarity of simulated networks affected network inference performance with fusion. The main factors governing similarity between our generated networks were the extent (and accuracy) of the orthology mapping, and the magnitude of variability between analogous interactions. We conducted a series of simulations to assess the effect of varying orthology coverage on networks in which analogous interactions varied little from one another. In order to determine the effect of the size of the orthology mapping on network inference performance, we simulated a series of 20 TFs by 200 genes networks at several values of percent orthology coverage.

As expected, increasing the weight of the fusion constraint $\lambda_S$ improved network recovery. Networks with higher orthology coverage saw a greater benefit from fusion, although the improvement was not limited to analogous interactions. One way of assessing the effect of fusion on network inference performance is to compute an ``exchange-rate'' between conditions in each network. Suppose you are interested in determining a regulatory network for species A, a model organism which is experimentally inconvenient. A related species, B, is easier to collect data for, but of less intrinsic interest. Our simulation results suggest that adding conditions from species B to a fused network inference problem has an effect on network A recovery that is similar to adding conditions from A at a reduced rate.

We also assessed the effect of varying the standard deviation of the distribution of differences between analogous interactions on network performance. FILL IN. MENTION PRIOR INTERPRETATION

\subsection{Fused regression improves performance on both the constrained and non-constrained parts of the network}
We wanted to know if using fused regression on networks with only conserved subgraphs affected the recovery of the non constrained parts of the networks. We used sets of two networks with 20 TFs and 200 genes each, where each set had 0\%, 25\%, 50\%, or 75\% of the TFs and genes in orthology groups. We varied lamS and computed AUPR using cross-validation on the portion of the network that had fusion constraints, the portion without any orthology information, and on the whole network. On the fused networks, performance improved as the fusion penalty weight increased; interestingly, performance gains are seen even the portion of the network where no orthology information is known. By constraining part of an underdetermined system, we obtain gains in both parts.

\subsection{SCAD successfully identifies and unfuses 'neofunctionalized' genes}
We recognize that orthology prediction is not necessarily a perfect proxy for functional conservation. We implemented an adaptive fusion algorithm that attempts to optimize a nonconvex saturating penalty function on differences between fused interactions. Pairs of interactions which are very dissimilar even after fusion, which sit in the flat portion of this penalty function, are effectively ``unfused''. We can interpret this ``unfusing'' as evidence for neofunctionalization of one or more of the genes involved in this pair of analogous interactions.

We performanced a simulation to assess the ability of our adaptive fusion algorithm to learn which parts of two input networks are conserved when orthology information is faulty. We generated synthetic fused networks and introduce error in the fusion constraints by adding false positives and negatives to the orthology information given to the solver. Because we know which entries in the orthology mapping are ``fake'' (not reflected in the generation of the networks), we can label fusion constraints that involve one or more ``fake'' mappings as being ``fake''. We verified that adaptive-fusion unfused mostly ``fake'' interactions, while leaving truly analogous interactions fused.

\subsection{SCAD improves network inference}
We wanted to determine whether SCAD is able to improve learning of GRNs. That is, beyond correctly identifying functional conservation and recovering the performance lost by fusing neofunctionalized genes, is our method able to produce more accurate networks than those learned without fusion?
